Designing a high-performance boundary element library with OpenCL and Numba

### 2021-04-07 opencl.tex
working through opencl section

To launch an OpenCL kernel the following steps need to be performed:
\begin{itemize}
	\item Initialize buffers on the compute device.
	\item Compute relevant data from host to the compute device.
	\item Load a kernel string and just-in-time compile it for the device.
	\item Execute the kernel for the given buffers and chosen index space.
	\item Copy results back to the host.
\end{itemize}
An excellent feature of OpenCL is its very good support for vectorized operations. OpenCL provides vector data types, such as \textbf{double4}, that can hold four double values in a SIMD register and defines a number of standard operations for these vector types. This makes it easy to explicitly target modern SIMD execution in a portable way while avoiding difficult compiler intrinsics and keeping kernel code readable.

Python has excellent OpenCL support through the PyOpenCL library by Andreas Kloeckner, which automates much of the initialization of the OpenCL environment and makes it easy to create buffers and launch OpenCL kernels from Python.

@@

Bempp-cl has OpenCL kernels for all its boundary operators. All operators have the same interface and are launched in the same way. In the first step the relevant data will need to be copied to the compute device. This data consists of:

\begin{itemize}
	\item Test and trial indices of triangles over which to be integrated.
	\item Signs of the normal directions for the spaces.
	\item Test and trial grid as flat floating point array, defining each triangle through nine floating point numbers, specifying the $(x, y, z)$ coordinates of each of the three nodes of a triangle.
	\item Test and trial connectivity information that stores the indices of each node for each triangle.
	\item Test and trial mappings of local triangle degrees of freedom to global degrees of freedom.
	\item Test and trial basis function multipliers for each triangle.
	\item Quadrature points and quadrature weights.
	\item A buffer that contains the global assembled matrix.
	\item Additional kernel parameters, such as the wavenumber for Helmholtz problems.
	\item Number of test and trial degrees of freedom.
	\item A single byte that is set to one if the test and trial grids are disjoint.
\end{itemize}
The actual kernel definition is shown in Figure \ref{fig:kernel_definition}.
Before the kernel can be launched, it needs to be configured and just-in-time compiled. Kernel configuration happens through C-style preprocessor definitions that are passed through the just-in-time compiler. These include the names of the test and trial space, the name of the function that evaluates the integral kernel, choosing between single and double precision types, and for SIMD enhanced kernel the vector length of the SIMD types.
For example, in Figure \ref{fig:kernel_definition} all floating point types have the name \ocl{REALTYPE}. This is substituted with either \textit{float} or \textit{double} during just-in-time compilation.

>>>>

To launch an OpenCL kernel the user must first initialize buffers on the compute device, then copy any relevant data from host to the compute device. A kernel string can then be loaded and just-in-time compiled for the device. The kernel is then run, and the results can be copied back to the host.

OpenCL has very good support for vectorized operations: it provides vector data types and defines a number of standard operations for these vector types. For example, the type \ocl{double4} will allow four double values to be held in a SIMD register. This makes it easy to explicitly target modern SIMD execution in a portable way while avoiding difficult compiler intrinsics and keeping kernel code readable.

Python has excellent OpenCL support through the PyOpenCL library by Andreas Kloeckner \cite{pyopencl}. PyOpenCL automates much of the initialization of the OpenCL environment and makes it easy to create buffers and launch OpenCL kernels from Python.

@@

Bempp-cl has OpenCL kernels for all its boundary operators. All operators have the same interface and are launched in the same way. In the first step, the relevant data will need to be copied to the compute device. This data consists of:

\begin{itemize}
	\item Test and trial indices of triangles over which to be integrated.
	\item Signs of the normal directions for the spaces.
	\item Test and trial grid as flat floating point array, defining each triangle through nine floating point numbers, specifying the $(x, y, z)$ coordinates of each of the three nodes of a triangle.
	\item Test and trial connectivity information that stores the indices of each node for each triangle.
	\item Test and trial mappings of local triangle degrees of freedom to global degrees of freedom.
	\item Test and trial basis function multipliers for each triangle.
	\item Quadrature points and quadrature weights.
	\item A buffer that contains the global assembled matrix.
	\item Additional kernel parameters, such as the wavenumber for Helmholtz problems.
	\item Number of test and trial degrees of freedom.
	\item A single byte that is set to one if the test and trial grids are disjoint.
\end{itemize}
The actual kernel definition is shown in \cref{fig:kernel_definition}.
Before the kernel can be launched, it needs to be configured and just-in-time compiled. Kernel configuration happens through C-style preprocessor definitions that are passed through the just-in-time compiler. These include the names of the test and trial space, the name of the function that evaluates the integral kernel, choosing between single and double precision types, and for SIMD enhanced kernel the vector length of the SIMD types.
For example, in Figure \ref{fig:kernel_definition} all floating point types have the name \ocl{REALTYPE}. This is substituted with either \textit{float} or \textit{double} during just-in-time compilation.
